Age-structured transmission dynamic models of RSV

### 2023-01-24 1_MSIS_model_explain/MSIS_model_explain.Rmd
update example

This age structured transmission dynamic model of RSV was first developed by Ginny Pitzer in her publication: https://doi.org/10.1371/journal.ppat.1004591. With the help of Dan Weinberger, we translated Ginny's Matlab codes to R codes. We further developed/modified this model, resulting in several mechanistic models corresponding to a variety of scenarios. Please refer to:

- https://doi.org/10.1001/jamanetworkopen.2021.41779
- https://doi.org/10.1038/s41541-022-00550-5
- https://doi.org/10.1101/2022.11.10.22282132

The complete code example for this tutorial please refer to
- https://github.com/weinbergerlab/RSV_metapop/tree/master/SimpleModel

@@

![Ginny's model](/Users/zhezheng/Library/CloudStorage/Box-Box/RSV_tutorials/RSV_tutorials/MSIS_model_explain/RSV_msis_model_structure.png)

@@

![Ginny's model](/Users/zhezheng/Library/CloudStorage/Box-Box/RSV_tutorials/RSV_tutorials/MSIS_model_explain/RSV_msis_model_structure.png)
\

@@

![Ginny's model](/Users/zhezheng/Library/CloudStorage/Box-Box/RSV_tutorials/RSV_tutorials/MSIS_model_explain/RSV_msis_model_structure.png)
::::

@@

Before we start, please learn the [Rstan basic](https://github.com/Gigi112/RSV_tutorials/tree/main/MSIS_model_explain/Introduction of Rstan.pptx) from the separated tutorials in the same file path. 
\
\
You will find three code documents inside code_source/stan_code file. age-structureMSIS_deterministic.stan is the stan code contains the ODE function, input data, input parameters, priors and the distribution of posterior distributions. ageMSIS_deterministric_annual.R is the R code that create the interface between stan and R to sample from posterior distribution. msis_wa_D.sh is a file to run the model on cluster. We will start with the stan code.
\
\
I suggest you first run the model in local for short iterations for debug purpose. If the stan model works, you can move to HPC by using the msis_wa_D.sh script.

>>>>

This age structured transmission dynamic model of RSV was first developed by Ginny Pitzer in her publication: [Environmental Drivers of the Spatiotemporal Dynamics of Respiratory Syncytial Virus in the United States](https://doi.org/10.1371/journal.ppat.1004591). With the help of Dan Weinberger, we translated Ginny's Matlab codes to R codes. We further developed/modified this model, resulting in several mechanistic models corresponding to a variety of scenarios. Please refer to:

- [Estimation of the Timing and Intensity of Reemergence of Respiratory Syncytial Virus Following the COVID-19 Pandemic in the U](https://doi.org/10.1001/jamanetworkopen.2021.41779)
- [Predicted effectiveness of vaccines and extended half-life monoclonal antibodies against RSV hospitalizations in children](https://doi.org/10.1038/s41541-022-00550-5)
- [Year-round RSV Transmission in the Netherlands Following the COVID-19 Pandemic - A Prospective Nationwide Observational and Modeling Study](https://doi.org/10.1101/2022.11.10.22282132)

@@

![Ginny's model](/Users/zhezheng/Library/CloudStorage/Box-Box/RSV_tutorials/RSV_tutorials/1_MSIS_model_explain/RSV_msis_model_structure.png)

@@

![Ginny's model](/Users/zhezheng/Library/CloudStorage/Box-Box/RSV_tutorials/RSV_tutorials/1_MSIS_model_explain/RSV_msis_model_structure.png)
\

@@

![Ginny's model](/Users/zhezheng/Library/CloudStorage/Box-Box/RSV_tutorials/RSV_tutorials/1_MSIS_model_explain/RSV_msis_model_structure.png)
::::

@@

Before we start, please learn the [Rstan basic](https://github.com/Gigi112/RSV_tutorials/tree/main/1_MSIS_model_explain/Introduction of Rstan.pptx) from the separated tutorials in the same file path. 
\
\
You will find three code documents inside code_source/stan_code file. age-structureMSIS_deterministic.stan is the stan code contains the ODE function, input data, input parameters, priors and the distribution of posterior distributions. ageMSIS_deterministric_annual.R is the R code that create the interface between stan and R to sample from posterior distribution. msis_wa_D.sh is a file to run the model on cluster. We will start with the stan code.
\
\
I suggest you first run the model in local for short iterations for debug purpose. If the stan model works, you can move to HPC by using the msis_wa_D.sh script.
